## Loading and preprocessing the data
```{r echo=TRUE}
data<-read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r echo=TRUE}
library(dplyr)
data$date<-as.Date(data$date)
TotalSteps<-data %>% filter(!is.na(steps)) %>% group_by(date) %>% summarize(sum=sum(steps,na.rm=TRUE))
TotalSteps
```

```{r echo=TRUE}
library(ggplot2)
ggplot(TotalSteps,aes(x=sum))+geom_histogram(binwidth = 1000)+labs(title="Steps Per Day",x="Total Steps",y="Frequency")
```

```{r echo=TRUE}
MeanSteps<-mean(TotalSteps$sum,na.rm=TRUE)
MeanSteps
```

```{r echo=TRUE}
MedianSteps<-median(TotalSteps$sum,na.rm=TRUE)
MedianSteps
```

## What is the average daily activity pattern?
```{r echo=TRUE}
AverageInterval<-data %>% group_by(interval) %>% filter(!is.na(steps)) %>% summarize(avg=mean(steps,na.rm=TRUE))

ggplot(AverageInterval,aes(x=interval,y=avg))+geom_line()+labs(title="Average daily steps",x="Interval",y="Average Steps")
```

```{r echo=TRUE}
MaximumInterval<-AverageInterval[which.max(AverageInterval$avg),]
MaximumInterval[[1]]
```

## Imputing missing values
```{r echo=TRUE}
sum(is.na(data))
```

##Impute missing values using mean and create new dataset dataNoNA
```{r echo=TRUE}
MissingData<-is.na(data$steps)
dataNoNA<-data
AvgInterval<-tapply(dataNoNA$steps,dataNoNA$interval,mean,na.rm=TRUE,simplify=TRUE)
dataNoNA$steps[MissingData]<-AvgInterval[as.character(dataNoNA$interval[MissingData])]
sum(is.na(dataNoNA))
```

```{r echo=TRUE}
TotalStepsNoNA<-dataNoNA %>% group_by(date) %>% summarize(sum=sum(steps))
ggplot(TotalStepsNoNA,aes(x=sum))+geom_histogram(binwidth=1000)+labs(title="Steps per day with missing values imputed",x="Interval",y="Frequency")
```

```{r echo=TRUE}
MeanStepsNoNA<-mean(TotalStepsNoNA$sum)
MeanStepsNoNA
```

```{r echo=TRUE}
MedianStepsNoNA<-median(TotalStepsNoNA$sum)
MedianStepsNoNA
```

##The impact of imputing missing data with the mean number of steps is both mean and median are equal (10766.19)

## Are there differences in activity patterns between weekdays and weekends?
```{r echo=TRUE}
dataNoNA<-dataNoNA %>% mutate(weekendOrWeekDay = ifelse(weekdays(date)=="Saturday" | weekdays(date)=="Sunday","Weekend","Weekday"))
```

##Plotting
```{r echo=TRUE}
AverageIntervalNoNA<-dataNoNA %>% group_by(interval,weekendOrWeekDay) %>% summarise(avg=mean(steps))

ggplot(AverageIntervalNoNA,aes(x=interval,y=avg,color=weekendOrWeekDay))+geom_line()+labs(title="Average steps per day by Weekend/Weekday",x="Interval",y="Steps")+facet_wrap(~weekendOrWeekDay,ncol=1,nrow=2)+theme(legend.position = "none")
```